---
title: "Exploring signed networks in Mastodon"
description: |
  An approach to empirically observe online signed networks, using Mastodon instance log data.
author:
  - name: Francisca Borquez
    url: https://orcid.org/0009-0009-7755-374X
    affiliation: VOSON Lab
    affiliation_url: http://vosonlab.net/
  - name: Bryan Gertzel
    url: https://orcid.org/0000-0002-6285-8649
  - name: Robert Ackland
    url: https://orcid.org/0000-0002-0008-1766
    
date: 2024-02-23
categories:
  - rstats
  - networks
  - rtoot
  - mastodon
  - signed networks
  - SNA
preview: friends-foes.png
output:
  distill::distill_article:
    self_contained: false
draft: false
bibliography: refs.bib
csl: apa-style.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

post <- basename(dirname(rstudioapi::getSourceEditorContext()$path))
data_dir <- file.path(Sys.getenv("VOSON_BLOG_DATA"), post, "/")
```

## 1. Introduction -- Exploring Mastodon networks with R tools

Our previous [post](https://vosonlab.github.io/posts/2023-07-27-mastodon-conversation-networks/) [@gertzel2023mastodon] introduced Mastodon -- a decentralised microblogging platform -- as well as code and methodological steps to collect Mastodon data using the `rtoot` R package [@SCHOCHCHAN2023] and construct networks for analysis. 

The present blog post uses `rtoot` to construct a signed network between Mastodon instances where a 'friend' tie indicates that an instance has nominated another instance as a peer while a 'foe' tie is where an instance blocks another instance. When an instance $i$ blocks another instance $j$ it means that content and users on instance $j$ are not visible to users on instance $i$.

While the present post uses `rtoot` directly, we have recently implemented Mastodon collection in `vosonSMl` and `vosonDash`, using the `rtoot` R package. Steps for collecting and generating different types of networks, following a similar framework as with the other data sources (Twitter, YouTube, Reddit, WWW) are available on the [GitHub page](https://github.com/vosonlab/vosonSML).

This blog post summarises research that was presented at the [Australian Social Network Analysis Conference ASNAC](https://www.ansna.org.au/asnac-2023), in November 2023. Research design, methodology and analysis by Francisca Borquez V.; data collection and analysis code by Bryan Gertzel and Rob Ackland.


## 2. Exploring signed networks in Mastodon
 
Mastodon is a decentralised microblogging platform that uses an open-source network protocol called `ActivityPub` to communicate between server instances and users. The instances collectively form a network of interoperable servers known as the federated universe or 'Fediverse'. Users join autonomous local communities (with their own rules, administration and moderation), typically established around a community or area of interest and intended to group similar users based on e.g. geographic location, language, views, interests, etc. 

The Mastodon federated governance of instances results in local rules and moderation, wherein users and other instances can be 'friended' but also silenced or suspended. As instances administrators apply moderating rules (e.g. by nominating foes), they protect their own users to be exposed to potentially harmful timelines, content and users. The following example provides an approach to empirically observe online signed networks, i.e. networks containing both positive (friendly) and negative (antagonistic) ties, using Mastodon instance log data.

### 2.1 Identifying seed server instances

Mastodon provides an [instance search engine](https://mastodon.help/instances), which is based on a database that gets crawled and updated on a daily basis. Users can define search parameters, for example by setting languages, descriptions, number of users, etc.

In this example, we are using 5 server instances relating to technology as starting points (seeds), we identified through the instance search engine. The seeds were qualitatively assessed and included in the sample if they:

- were self-identified and 'actually' related to tech
- were actively moderated and have terms of use in place
- presented a list of moderated servers or instance blocks (foes)
- had at least 1,000 users, who are involved and active

| Instance | Number of users |
| ----------- | ----------- |
| social.veraciousnetwork.com | 2,087 |
| defcon.social | 1,122 |
| vmst.io | 1,942 |
| social.linux.pizza | 1,675 |
| gamestoot.de | 1,798 |

Table 1: seed server instances related to tech and number of active users.

Data were collected in November 2023. We used the R package `rtoot` [@SCHOCHCHAN2023] to programmatically access the list of ‘friends’ -- federated servers, with the `get_instance_peers` function -- and ‘foes’ -- moderated servers, with the `get_instance_blocks` function.  The following code conducts the data collection and saves the collected raw data as an RDS file, for later use. 

```{r, eval=FALSE, echo=TRUE}
# Code authored by Bryan Gertzel, VOSON Lab
# options
options(scipen = 999)
options(encoding = "UTF-8")

library(tidyverse)
library(rtoot)

#--------------------
#Collect the raw data
#--------------------

# read from csv. 'About' column contains server instances URLs. Trailing gets removed with second line.
seeds <- read_csv(file = "seeds.csv") |> pull("About") |> str_remove_all("https://|/about|/explore")

# get instance peer list (friends)
get_peers <- function(x) {
  tryCatch({
    peers <- get_instance_peers(x, anonymous = TRUE)
    tibble(peers.domain = peers) |> mutate(instance = x) |> relocate(instance)
  },
  error = function(e) {
    message(paste(x, "-", "get_peers", e))
    NULL
  })
}

# get instance blocks list
get_blocks <- function(x) {
  tryCatch({
    blocks <- get_instance_blocks(x, anonymous = TRUE)
    blocks |> rename_with(~ paste0("blocks.", .x)) |> mutate(instance = x) |> relocate(instance)
  },
  error = function(e) {
    message(paste(x, "-", "get_blocks", e))
    NULL
  })
}

# get friends and foes for seed instances in a dataframe
data <- map_dfr(seeds, function(x) bind_rows(get_peers(x), get_blocks(x)))

nrow(data)
#[1] 223032

#Save raw data
saveRDS(data, "data_23Nov.rds")
```

### 2.2 Constructing the full network and categorising nodes
The resulting dataframe contains 223,032 rows, including both peers and foes nominations. We then constructed a directed network, where nodes are server instances, and there are two types of tie: ‘friend’ i.e. where instance $i$ regards instance $j$ as a ‘peer’, and ‘foe’ i.e. where instance $i$ blocks instance $j$. Type of tie -- friend or foe -- was included in the network as edge attribute. 

```{r, eval=TRUE, echo=TRUE}
# Code authored by Bryan Gertzel and Rob Ackland, VOSON Lab
#Construct the networks
library(dplyr)
library(igraph)

data <- readRDS(paste0(data_dir, "data_23Nov.rds"))

nrow(data)

# server relations (edge list)
relations <- data |>
  select(from = instance, to = peers.domain) |>
  filter(!is.na(to)) |>
  mutate(type = "friend") |>
  bind_rows(
    data |>
      select(from = instance, to = blocks.domain) |>
      filter(!is.na(to)) |>
      mutate(type = "foe")
  ) |>
  distinct()

g <- graph_from_data_frame(relations)

vcount(g)
ecount(g)

# Option for signed networks -- weighted ties 
#E(g)$weight <- ifelse(E(g)$type=="friend", 1, -1)

#identifying seed servers
V(g)$seed <- "no"
V(g)$seed[which(degree(g, mode="out")>0)] <- "yes"      

table(E(g)$type)
```

The full network contains 124,258 nodes (server instances) and 223,032 ties, of which 222,389 are positive ('friend' or peer nominations) and 643 are negative ('foe' nominations).

Then, the nodes were classified as 'friends' , 'foes', 'mixed' and 'neither', according to the type of ties they receive:

| Type | Classification | Number of nodes | Colour |
| ----------- | ----------- | ----------- |----------- |
| Friend | if receive >=2 friend nominations and <2 foe nominations | 28,036  | Green | 
| Foe | if receive <2 friend nominations and >=2 foe nominations | 41  | Red | 
| Mixed | if receive >=2 friend nominations and >=2 foe nominations | 52 | Orange | 
| Neither | Otherwise | 96,129  | White | 

Table 2: node classification according to type of tie.

The following code classifies the nodes and saves the `igraph` graph as a `graphml` file, for further analysis.

```{r, eval=FALSE, echo=TRUE}
# Code authored by Rob Ackland, VOSON Lab

#-----------------------------
#Classify nodes: 
#if receive >=2 friend nominations and <2 foe nominations then "friend"
#if receive <2  friend nominations and >=2 foe nominations then "foe"
#if receive >=2 friend nominations and >=2 foe nominations then "mixed"
#otherwise: "neither"
#Note: this code takes several minutes to run

e_ind <- incident_edges(g, V(g), mode="in")

#e_ind[1]

f1 <- function(t){
 
    x1 <- table(e_ind[[t]]$type)
    isFriend <- 0
    if ("friend" %in% names(x1))
        if (x1[which(names(x1)=="friend")]>=2)
            isFriend <- 1
    isFoe <- 0
    if ("foe" %in% names(x1))
        if (x1[which(names(x1)=="foe")]>=2)
            isFoe <- 1
    
    if (t%%100==0)
        cat("finished:", t, "\n")

    type <- ifelse(isFriend & !isFoe, "friend",
                   ifelse(!isFriend & isFoe, "foe", 
                          ifelse(isFriend & isFoe, "mixed", "neither")))
}

#This takes several minutes
#L <- lapply(1000:2000, f1)                #testing
L <- lapply(1:length(e_ind), f1)
df <- do.call("rbind", L)

V(g)$type <- df

table(V(g)$type)
#foe  friend   mixed neither 
#41   28036      52   96129

#save graphml
write.graph(g, "g.graphml", format="graphml")  
```

### 2.3 Constructing a subnetwork of negative ties only

Then, we constructed a subnetwork that includes 'foe' ties only. The nodes in this subnetwork encompass the seed instances which have sent negative ties and those instances that have received at least one 'foe' nomination. The resulting network has 497 nodes and 643 'foe' edges. Isolates were removed. Nodes were colour coded based on the categorisation. 

```{r, eval=TRUE, echo=TRUE}
# Code authored by Rob Ackland, VOSON Lab

#Read full graph
g <- read.graph(paste0(data_dir,"g.graphml"),format="graphml")

#Construct network: (1) only foe ties and (2) remove isolates
#So this network only contains seeds and nodes that have received at least one foe nomination

g5 <- delete.edges(g, which(E(g)$type=="friend"))
g5 <- induced.subgraph(g5, which(degree(g5)>0))

#Colour nodes according to their status as seed/friend/foe/mixed
#So we find that there are nodes in this network that are coloured green: they are 
#classifiedas "friend" (more than 2 friendship nominations) but at least one seed has classified them as foe

V(g5)$color <- "white"
V(g5)$color[which(V(g5)$type=="friend")] <- "green"       #seed servers
V(g5)$color[which(V(g5)$type=="foe")] <- "red"       #seed servers
V(g5)$color[which(V(g5)$type=="mixed")] <- "orange"       #seed servers
V(g5)$color[which(V(g5)$seed=="yes")] <- "blue"       #seed servers

#png("foe_ties_only.png", width=600, height=600)
#plot(g5, vertex.size=3, vertex.label="", edge.arrow.size=0.3)
#dev.off()

write.graph(g5, "g5.graphml", format="graphml")    

table(V(g5)$type)

#Let's just check that a node classified as mixed really is mixed
mixed <- which(V(g5)$type=="mixed")
#Example:
V(g5)$name[mixed[1]]

#Let's see the inbound ties to this node
e_ind <- incident_edges(g, V(g), mode="in")

e_ind[[which(V(g)$name=="iddqd.social")]]

#Let's see the type of tie
e_ind[[which(V(g)$name=="iddqd.social")]]$type

#So everything looks correct: this node is "mixed" as there are two friendship nominations
#and 4 foe nominations

```

The `graphml` file was then read into Gephi and the following visualisation was produced.

![Figure 1: Mastodon network of server instances. Seed instances are coloured blue. Nodes (other than seeds) represent Mastodon instances which have received at least one foe nomination by a seed and are classified by colour: green – ‘friend’; orange – ‘mixed’: red – ‘foe’, and white – ‘neither’. Node size by indegree. Isolates were removed from the network. Edge reflect incident node colour. Network visualisation produced with Gephi.](Network-negative ties and seeds.png)


## 3. Findings

Consistent with the literature [@EVERETT2014111; @STADTFELD2020129], the network structure presents group boundaries around the seed servers, given that instance moderators are consciously separating themselves from those instances that have been identified as unequivocal ‘foes’(red) or ‘mixed’ (orange).

In their exploration of online signed networks, [@leskovec2010signed] sustained that positive ties are expected to produce clusters while negative ties tend to span positive clusters, which can be observed in this example. In this network, green nodes (friends) – those servers that have received only 1 negative nomination and 2 or more positive nominations by other seed instance - tend to cluster around the seed servers, while unequivocal ‘foe’ servers (red) plus ‘mixed’ servers (orange)- which have received at least 2 negative and positive nominations - are localised in a central area of the network, spanning the clusters.


## 4. Next Steps
The next steps of this research will involve the use of `signnet` [@Schoch2023], an R package that provides methods to analyse signed networks, with special focus on structural balance using triads, and signed blockmodeling. 

Another approach will involve to qualitatively assess seed servers' moderating rules, based on what constitutes an 'acceptable' reason to block an instance. Moderating rules will be then categorised in scales, from ‘strict’ to ‘permissible’, and such values can be included in the network as node attributes. Similarly, we will assess the capacity of the instance administrator(s) to ‘defend their space’ as a network strategy to define boundaries, i.e. by identifying other variables that could explain clustering, for example, frequency of activity (active versus passive moderation), date of creation (established communities versus new entrants), etc. 

