---
title: "Egocentric Networks from Twitter timelines"
description: |
  Demonstration of how to use rtweet and vosonSML to construct an ego net from Twitter users timelines.
author:
  - name: Bryan Gertzel
    affiliation: VOSON Lab
    affiliation_url: http://vosonlab.net/
date: 2022-06-05
categories:
  - rstats
  - twitter
  - networks
  - egocentric
  - rtweet
  - vosonsml
  - timeline
preview: egocentric_400x400.png
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 3
draft: false
bibliography: refs.bib
csl: apa-style.csl
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
data_path <- "E:/My Drive/VOSON-BLOG-DATA/2022-06-05-egocentric-networks-from-twitter-timelines/"

library(dplyr)
library(ggraph)
library(igraph)
library(rtweet)
library(vosonSML)
```

## Introduction

Egocentric networks or ego nets are networks that focus on a particular actor (the ego) and map out their connections to other actors. In an ego net other actors are referred to as alters, and by collecting the outward expanding connections of alters, ego nets of varying degrees can be constructed [see @RN59, p.168]. Some literature and software refer to these as neighborhood networks, with varying orders instead of degrees. For example in a friendship network, a neightborhood network of the first order (1.0 degree) contains just the friends of the ego, whereas a network of the second order (sometimes second step) also contains "friends of friends" (a 2.0 degree ego net).

By collecting the tweets in a Twitter users timeline, and the timelines of users referenced, we can create a 1.0, 2.0 or 1.5 degree network for the ego. A 1.5 degree network is similar to the 1.0 degree, except it also contains relationships or ties between the alters, or "between friends" of the ego from the previous friendship network example.

It should be noted that by using user timelines that this is not necessarily a friendship network, but instead a network of twitter users who are associated through tweet activity. This kind of ego net can lead to insights beyond simply declared friendships (obtained from Twitter's friend/follower metadata) as the network structure is the result of users interactions on the platform over a period of time.

This post will demonstrate how to construct an ego net from a twitter users timeline using the `rtweet` package to collect tweets and `vosonSML` to create an actor network.

## Collect the ego timeline

The first step is to collect the ego users timelime. In this post we will be using the VOSON Lab `@vosonlab` twitter account, and collecting the Twitter timeline using `rtweet`. The Twitter API restricts the number of timeline tweets that can be collected to the [most recent 3,200 tweets](https://developer.twitter.com/en/docs/twitter-api/tweets/timelines/introduction), but we can set this to a lesser value e.g most recent 100 tweets, and also use the same parameter for alters timelines for the purposes of this demonstration.

*Please note this post uses vosonSML v0.30.6 and rtweet 0.7.0*

```{r ego, echo = TRUE, eval = FALSE}
library(dplyr)
library(DT)
library(rtweet)
library(vosonSML)

ego_user <- "vosonlab"

# get twitter user timeline
ego_tweets <- get_timeline(c("vosonlab"), n = 100, token = NULL)

# get twitter user id
ego_user_id <- slice_head(ego_tweets |> count(user_id, sort = TRUE)) |>
  pull(user_id)

# create actor network from timeline tweets
ego_net <- ego_tweets |> ImportData("twitter") |> Create("actor")
```

```{r ego-read-data, include = FALSE}
ego_tweets <- readRDS(paste0(data_path, "vosonlab_timeline_n100.rds"))

ego_user_id <- slice_head(ego_tweets |> count(user_id, sort = TRUE)) |>
  pull(user_id)
# ego_net <- ego_tweets |> ImportData("twitter") |> Create("actor")
```

```{r ego-net-output}
ego_net <- ego_tweets |> ImportData("twitter") |> Create("actor")
```

A result of **41 nodes** indicates that there are **40 alters** in the network with ties to the ego. 

```{r degree-1, echo = FALSE, fig.cap = "vosonlab 1.0 degree actor network"}
net_deg_1 <- ego_net
net_deg_1$nodes <- net_deg_1$nodes |>
  mutate(node_cat = ifelse(user_id %in% ego_user_id, "ego", "alter")) |>
  mutate(size = ifelse(node_cat == "ego", 3, 1))

node_cat_colors <- list(ego = "magenta", alter = "blue")

g_deg_1 <-
  graph_from_data_frame(net_deg_1$edges, vertices = net_deg_1$nodes)
V(g_deg_1)$degree <- degree(g_deg_1)

set.seed(100)
layout <- create_layout(g_deg_1, layout = "star")
p_deg_1 <- ggraph(layout) +
  geom_edge_fan(colour = "gray") +
  geom_node_point(aes(colour = factor(node_cat), size = size)) +
  scale_color_manual(values = node_cat_colors) +
  labs(color = "nodes") +
  guides(size = "none")

p_deg_1
```

## Collect timelines of alters

From the previous step we created an actor network represented as nodes and edges dataframes from the ego users timeline. We can now use this to extract all of the user ids of the alters in the network.

Note that we haven't specified the degree of the ego net at this stage, however by virtue of the twitter data (timeline tweets) having all been created by the ego user, we can assume all of the alters (referenced users) are connected to the ego in this network.

```{r alters, echo = TRUE, eval = FALSE}
# get list of alter user ids
alter_user_ids <- ego_net$nodes |>
  filter(user_id != ego_user_id) |> distinct(user_id) |> pull()

# get 100 most recent tweets from all of the alters timelines
alter_tweets <- get_timeline(alter_user_ids, n = 100, token = NULL)
```

```{r alters-1a, include = FALSE}
alter_user_ids <- ego_net$nodes |> filter(user_id != ego_user_id) |>
  distinct(user_id) |> pull()

alter_tweets <- readRDS(paste0(data_path, "vosonlab_alters_timeline_n100.rds"))

```

```{r alters-1b, echo = FALSE}
print(alter_tweets, n = 3)
```

## Create an actor network

Now that all of the tweets from the alters timelines have also been collected, the data can be merged and a single actor network created. This actor network can be considered a **2.0 degree** network, as it contains not only the associations or "friends" from the egos timeline, but also the associations or "friends" of the alters from their timelines.

```{r alters-2, echo = TRUE, eval = FALSE}
# combine all of the tweets from ego and alters timelines
tweets <- bind_rows(ego_tweets, alter_tweets)

# create actor network from combined timeline tweets
actor_net <- tweets |> ImportData("twitter") |> Create("actor")
```

```{r import-comb-actor, echo = FALSE}
tweets <- bind_rows(ego_tweets, alter_tweets)
actor_net <- tweets |> ImportData("twitter") |> Create("actor")
```

Here we can see an actor network of **2120 nodes** and **6036 edges**, substantially larger than our initial actor network.

```{r degree-2, echo = FALSE, fig.cap = "vosonlab 2.0 degree actor network"}
net_deg_2 <- actor_net
net_deg_2$nodes <- net_deg_2$nodes |>
  mutate(node_cat = ifelse(
    user_id %in% ego_user_id,
    "ego",
    ifelse(user_id %in% alter_user_ids, "alter", "alter-alter")
  )) |>
  mutate(size = ifelse(node_cat == "ego", 4, ifelse(node_cat == "alter", 1.2, 1)))

net_deg_2$edges <- net_deg_2$edges |>
  mutate(edge_type = ifelse(
    stringr::str_detect(edge_type, "mention"),
    "mention",
    edge_type
  ))

node_cat_colors <-
  list(ego = "magenta",
       alter = "blue",
       `alter-alter` = "darkgray")

g_deg_2 <-
  graph_from_data_frame(net_deg_2$edges, vertices = net_deg_2$nodes)
V(g_deg_2)$degree <- degree(g_deg_2)

set.seed(100)
layout <- create_layout(g_deg_2, layout = "fr") #, root = ego_user_id)
p_deg_2 <- ggraph(layout) +
  geom_edge_link(colour = "gray") +
  geom_node_point(aes(colour = factor(node_cat), size = size)) +
  scale_color_manual(values = node_cat_colors) +
  labs(color = "nodes") +
  guides(size = "none")

p_deg_2
```

## Optionally add user metadata as node attributes

At this point we can optionally add some user metadata to our network as node attributes. This allows us to change visual properties of the network graph based on actor attributes. For example, we could map the node size to number of followers a twitter user may have.

Please note this step requires a [vosonSML twitter auth object](https://vosonlab.github.io/vosonSML/reference/Authenticate.twitter.html) if you want to use the look up feature for complete users metadata.

```{r comb-add-meta, eval = FALSE, echo = TRUE}
# this step requires a previously saved vosonSMML twitter auth object
twitter_auth <- readRDS("~/.vsml_oauth1a")

# add user profile metadata
actor_net_meta <- actor_net |>
  AddUserData(tweets, lookupUsers = TRUE, twitterAuth = twitter_auth)
```

```{r comb-add-meta-file, include = FALSE}
# saveRDS(actor_net_meta, paste0(data_path, "vosonlab_actor_net_meta.rds"))
actor_net_meta <- readRDS(paste0(data_path, "vosonlab_actor_net_meta.rds"))
```

Here is some of the actor metadata available and an example of how it can be presented and explored using a data table.

```{r comb, echo = TRUE}
# node attributes
names(actor_net_meta$nodes)

# explore actors metadata
actors_table <- actor_net_meta$nodes |>
  filter(user_id %in% c(ego_user_id, alter_user_ids)) |>
  mutate(screen_name = paste0("@", screen_name)) |>
  select(name = screen_name, display = display_name, location, description,
         followers = followers_count, tweets = statuses_count) |>
  slice_head(n = 5)

library(reactable)

reactable(actors_table, bordered = TRUE, striped = TRUE, resizable = TRUE,
          wrap = FALSE, searchable = TRUE, paginationType = "simple")
```

## Create an ego subgraph

A 1.5 degree network can be useful to reveal the associations between an egos alters. This can be achieved by creating a subgraph of the 2.0 ego network that retains only the previously identified alters (see below `igraph::induced_subgraph`). As we know every alter is connected to the ego so it is also often useful to visualise ego nets without the ego as it is then easier to see clustering.

```{r graph, echo = TRUE}
library(igraph)

# use the vosonSML to convert the network dataframes into an igraph object
g <- actor_net_meta |> Graph()

# create a subgraph with ego removed
g2 <- induced_subgraph(g, c(alter_user_ids))

g2
```

As we saw in our initial actor network constructed from only the ego users timeline we now have **40 nodes** again, matching the number of alters. This actor network has many more edges however, as **1862 ties** or relations between the alters were captured from the collection of the alters timelines. 

## Visualise with an interactive visNetwork plot

Using the `igraph` and `visNetwork` package we can create a simplified and undirected ego network graph of alters. Community detection can be performed and visualised using the `igraph` walktrap clustering algorithm and Fruchterman-Reingold force-directed layout. We can further map some visual properties of nodes to attributes - with node size corresponding to the node degree, edge width to combined weight, and color to clustering community group.

```{r net, echo = TRUE, fig.cap = "vosonlab 1.5 degree actor network"}
library(visNetwork)

# combine and weight the edges between nodes
E(g2)$weight <- 1
g2 <- simplify(g2, edge.attr.comb = list(weight = "sum"))
g2 <- as.undirected(g2)

# perform some community detection using a random walk algorithm
c <- walktrap.community(g2)
V(g2)$group <- membership(c)

# map visual properties of graph to attributes
E(g2)$width <- ifelse(E(g2)$weight > 1, log(E(g2)$weight) + 1, 1.1)

V(g2)$size <- degree(g2) + 5
V(g2)$label <- paste0("@", V(g2)$screen_name)

visIgraph(g2, idToLabel = FALSE) |>
  visIgraphLayout(layout = "layout_with_fr") |>
  visOptions(
    nodesIdSelection = TRUE,
    highlightNearest = TRUE
  )
```


The final result is a ego net with some clear associations. Colours and placement of nodes found to represent some interesting domains and community relationships between the Twitter `@vosonlab` account and its timeline network alters. Isolates representing more distant connections with no detected community ties.
